### 2025-02-25 paper/conclusions.tex
Rewrite conclusions.

The paper demonstrates a technique of developing high performance applications with
strong correctness guarantees.  The key insight lies in using a proof assistant
in cooperation with a high-performance language of choice.  This gives a clear
separation of concerns that is very difficult to achieve within a single
language.  The proof assistant is used to design a specification, prove
all the correctness invariants of interest and performs an extraction
into a high-performance language of choice.  This may take a non-trivial
effort, but correctness \emph{is} demanding!

Having a trusted specification as well as entire code-generation pipeline
within a single dependently-typed framework is incredibly powerful.
As we have demonstrated at the example of the neural network, we can introduce domain-specific
optimisations and even swap the backend in case its performance is unsatisfying.
For our example, the entire framework that includes array theory, DSL,
optimisations and extraction is about 2000 lines of Agda code.  We managed
to introduce a new backend in about two days and match performance of
the hand-written code.

A lot of pieces that we have developed in this
paper can be reused in other numerical applications.  We used dependent
types to guarantee the absence of out-of-bound indexing, certain function being
inverses as well as well-scopedness and well-typedness of our DSL.
However, there are many more opportunities that we did not explore.
For example, one can prove the correctness of optimisations, relating
evaluation of optimised an non-optimised expressions.  We can provide
more guarantees when we run extraction, \eg{} we can formalise some
aspects of the backend language and relate them to our DSL.
As for the DSL itself, we can try extending it with internal let
constructions which should improve our C code generation as well as
facilitate optimisations of the derivatives.

There are indeed plenty of opportunities, but the key point is this.
Correctness and performance are competing requirements when it
comes to application design.  Therefore, such a cooperation between
correctness-oriented and performance-oriented tools is likely to
persist.  With this work we demonstrate that such cooperation
is feasible in practice.

>>>>

The paper demonstrates a technique of developing high performance
applications with strong correctness guarantees, including the absence
of out-of-bound indexing, certain functions being inverses, as well as
well-scopedness and well-typedness of our DSL.

The key insight lies in using a proof assistant in cooperation with a
high-performance language of choice. This gives a clear separation of
concerns that is very difficult to achieve within a single language.
The proof assistant is used to design a specification, prove all the
correctness invariants of interest and performs an extraction into a
high-performance language, in our case Futhark, although only a
relatively small part of our work is specific to that language.

Having a trusted specification as well as entire code-generation
pipeline within a single dependently-typed framework is incredibly
powerful. As we have demonstrated at the example of the neural
network, we can introduce domain-specific optimisations and
transformations, such as automatic differentiation. For our example,
the entire framework that includes array theory, DSL, optimisations
and extraction is about 2000 lines of Agda code.

A lot of pieces that we have developed in this paper can be reused in
other numerical applications. However, there are many more
opportunities that we did not explore. For example, one can prove the
correctness of optimisations, relating evaluation of optimised an
non-optimised expressions. We can provide more guarantees when we run
extraction, \eg{} we can formalise some aspects of the backend
language and relate them to our DSL. As for the DSL itself, the exact
choice of supported primitives, and its implications for extracting
high performance code or performing optimisation or transformation,
remains an interesting question.

There are indeed plenty of opportunities, but the key point is this.
Correctness and performance are competing requirements when it comes
to application design. Therefore, such a cooperation between
correctness-oriented and performance-oriented tools is likely to
persist. With this work we demonstrate that such cooperation can be
done using fairly straightforward means, by Agda standards, and obtain
compelling practical performance.
